---
title: "Plot Weights & Locations"
author: "steppe"
output:
  word_document: default
always_allow_html: true
---


This AIM study was designed with three major principals in mind. The first, was maximizing the number of environments which it represents across the field office. The UFO field office varies by nearly XXXX feet, XXXX average precipitation, XXXX average temperatures, and displays a diversity of environments; and this sample design included this entire range of variation. To ensure that these diverse areas were represented, ten major strata were identified *see table*, in order to ensure that they had plots located in them in order to evaluate their ecological *integrity*. However, the management concerns, actions, and heterogeneity (internal dissimilarity) of these strata differ. To best understand the ecological context of these strata, and what management actions the BLM may take, the proportion of plots were tailored to each stratum. The third principal was that all plots be randomly placed within the stratum, which allows *inference* to be made from plots to the, unsampled, entirety of the stratum. This *weighted sample design* allows us to maximize statistical understanding while minimizing the field effort required to undertake it. 

The initial sample design followed the general AIM implementation design to contain 255 plots to be sampled across five years. Sampling over a five year period is essential in order to ensure that plots are visited during periods wherein the vegetation is *reproductively active*, hence identifiable. A benefit of this prolonged time frame is that anomalous weather conditions are unlikely to affect the plots across the entire time period. For example, the condition of plots may be compared in a wet year, to a dry year, as referenced from a year with typical rainfall. In order to avoid random unexpected processes which cluster in space which may occur over the period of conducting the sample design, the entire design is split into a *panel* for each year of sampling. Each panel is composed of a subset of randomly selected points within the sample design and avoids missing swathes of the *target frame* during the sample period due to an event such as a prolonged wildfire. 

The location of the random plots for AIM are generated via computer and a number of them may not feasibly be sampled. In general potential plots are most often, anecdotally, rejected for: being located on steep slopes which are dangerous for field crews to sample, requiring access to cross private land to access BLM which landowners deny. The inability of these plots to be sampled reduces the spatial extent of the strata which we may use statistics to infer across, and increases the measures of uncertainty associated with these areas. In this section we review the original sample design, derive original sample weights for sub-units of the field office, and calculate the weights for each of these units after the sample design has been completed. 

#### the design may be represented in simple mathematical terms

Using an example from a stratum which is of high land management importance, sagebrush-steppe, we illustrate the site selection process. While the aerial extent of sagebrush-steppe in the target frame is roughly one quarter (0.246), the stratum makes up one third (0.33) of all plots

In order to convey the number of plots drawn in any stratum the following equation is representative. 

$$
n \;=\; \dfrac{N * \pi_{i}} {1 / panels}
$$

+ $N$ the total number of plots in the sample design, *i.e. 255*
+ $\pi_{i}$ the inclusion probability of a plot in each stratum being drawn, *e.g. 0.33 targets the placement of a third of all plots to be put into the Sage-steppe Stratum*
+ $panels$ the number of design panels, *i.e. the number of years to stretch sampling across*

For sagebrush steppe we will then have: 

$$ 
16.83 \;=\; \dfrac{255 * 0.33 }{1/5}  
$$
plots for each panel, which will round down to 16 in order to accommodate representation of some of our other strata.


The **weight** of a plot is the amount of acres it represents within a stratum, we utilize this metric to help derive *measures of uncertainty* while making inference from our sampled plots to the whole stratum. 

$$
W_{i} = \dfrac{Stratum_{area}}{n}
$$

+ ${W_{i}}$ the weight acres, *i.e. the area which a plot represents within the target frame*
+ ${Stratum_{area}}$ the total area of the stratum in the area of analysis *e.g. 214,023 acres of sage-steppe*
+ ${n}$ the total number of plots in a stratum *e.g. for a panel of sage steppe 16*

Each sagebrush-steppe plot will have a weight of

$$
13,376 \;=\; \dfrac{214,023}{16}
$$
acres per plot, which it represents. 

#### Areas of Analysis

The original AIM design covers the entire target frame of the UFO field office. However, a few additional analytical  and management sub-units exist within this extent. Their are two National Monuments - the Gunnison Gorge & Dominguez-Escalente - the latter of which is partially administered by the Grand Junction Field Office (GJFO), which have different management directives which distinguish them from the majority of land in the target frame. Likewise there are numerous Wilderness Study Areas (WSA's), and Area's of critical Environmental Concern (ACEC's), throughout the field office. The UFO intends these areas to have higher proportions of certain metrics relative to the remainder of BLM land. 

> "*Maximize native vegetation and natural processes by ensuring upland vegetation communities are within the range of natural variability, with an appropriate mix of plant functional groups, cover, and diversity, according to best available science ... over 10 years with 80 percent confidence*" 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(dpi = 300) 
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
```

```{r Load libraries, warning = F, message = F, echo = F}
library(tidyverse)
library(sf)
library(terra)
library(spsurvey)
library(ggalluvial)
source('functions.R')
```

```{r}
p2d <- '../data/raw'

plot_status <- read.csv(file.path(p2d, list.files(path = p2d, pattern = "summary.csv")))  %>% 
  mutate(across(.cols = everything(), ~ str_to_lower(.))) %>% 
  
  # humans are humans and typos are an expression of humanity. also  add a 
  # factor levels which we arrange by and slice the top from . 
  mutate(
    Plot.Status = str_trim(Plot.Status, side = "both"),
  #  Plot.Status = str_replace(Plot.Status, 'rejected', 'not_sampled'),
    Plot.Status = str_replace(Plot.Status, " ", "_"),
    psF = factor(Plot.Status, levels = c('sampled', 'not_sampled', 'rejected'))
    ) %>%  
  
  # some numbers are missing leading zereos
  separate(Plot.ID, into = c('stratum', 'id'), sep = '-') %>% 
  mutate(id = str_pad(id, 3, pad = "0")) %>% 
  
  # group up - and take that 'sampled' from the top - if present
  unite('Plot.ID', stratum:id, sep = "-", remove = F) %>% 
  group_by(Plot.ID) %>% 
  arrange(psF) 

# some plots were not sampled in one year, but subsequently sampled
status_check <- plot_status %>% # we want to ensure all of these plots are returned as sampled. 
  mutate(grp_n = n()) %>% 
  filter(grp_n >= 2, Plot.Status == 'sampled')

plot_status <- plot_status %>% 
  slice_head(n = 1) %>% 
  ungroup()

# inner_join(plot_status, status_check, by = 'Plot.ID') # take a look for yourself, all good. 

# it turns out some oversamples from the original design were not populated over to the yearly tracking 
# sheets. We can note there status as not_sampled here. 
plot_status <- plot_status %>% 
  mutate(across(.cols = Plot.ID:stratum, ~ str_to_upper(.x)))
pts <- st_read(
  '/media/sagesteppe/ExternalHD/aimDB/data/raw/AIM_Sample_Design/AIM_Design_Stratification.shp',
                quiet = T) %>% 
  left_join(., plot_status, by = c('PLOTID' = 'Plot.ID')) %>% 
  select(-Panel) %>%  
  separate(PLOTID, into = c('stratum', 'id'), sep = '-', remove = F) %>% 
  mutate(across(Plot.Status:psF, ~ replace_na(.x, 'not_sampled'))) %>%  
  # some over samples were not completed
  select(Plot.ID = PLOTID, stratum, id, Panel = PANEL, Plot.Status, psF, xcoord, ycoord, geometry)
  
# st_write(pts, 
#          '/media/sagesteppe/ExternalHD/UFO_cartography/AIM_plots_outcomes/plots_outcomes.shp',
#          quiet = T)

plot_status <- pts %>% 
  select(colnames(plot_status)) %>% 
  st_drop_geometry()

plot_summary <- plot_status %>% 
  group_by(stratum, Plot.Status) %>% 
  summarise(n = n())  %>% 
  ungroup(Plot.Status) %>% 
  mutate(total = sum(n)) %>% 
  pivot_longer(n:total, values_to = 'Number') %>% 
  mutate(Plot.Status = if_else(name == 'total', 'total', Plot.Status)) %>% 
  select(-name) %>% 
  group_by(stratum, Plot.Status) %>% 
  sample_n(1)

rm(status_check)

# format table for printing
plot_summary <- plot_summary %>% 
  pivot_wider(values_from = Number, names_from = Plot.Status) %>% 
  mutate(stratum = str_to_upper(stratum)) %>% 
  mutate(across(.cols = everything(), ~ replace_na(.x, 0))) %>% 
  rename_with(., ~ str_to_sentence(.))

# DERIVE THE VALUES FOR BOTH TOTAL AREA AND AREAS IN ACRES
r <- read.csv(file.path(p2d, 'UFO_strata_areas.csv')) %>% 
  mutate(ProportionalArea = Cells/sum(Cells)) %>% 
  select(Stratum, Code, ProportionalArea, Acres)

deSS <- data.frame( # desired sample size per stratum over live of sample design
  'Stratum' = c("AS", "GR", "MC", "MMS", "OT" , "PJ", "PP", "RI", "SD" , "SS" ), 
  'DesiredSS' = c(5, 5, 15, 25, 5, 25, 5, 15, 75, 80),
  'PropTarget' = c(0.01, 0.02, 0.05, 0.1, 0.01, 0.12, 0.01, 0.05, 0.3, 0.33)
) 

OriginalWeights <- left_join(plot_summary, r, by = c('Stratum' = 'Code')) %>% 
  left_join(., deSS, by = 'Stratum') %>% 
  select(StratumName = Stratum.y, Stratum, Total,  NotSampled = Not_sampled, Rejected, Sampled, 
          DesiredSS, PropArea = ProportionalArea, PropTarget, Acres) %>% 
  mutate(ApproxStWgt = Acres / (DesiredSS/5))

OriginalWeights %>% 
  mutate(PropArea = round(PropArea, 3)) %>% 
  mutate(across(.cols = c(Acres, ApproxStWgt),  round)) %>% 
  arrange(-Acres) %>% 
  ungroup() %>% 
  select(Stratum, Acres, PropArea, PropTarget, DesiredSS, ApproxStWgt) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","), 
               col.names = c('Stratum', 'Total Area (acres)', 'Prop. Area',
                             'Prop. Site', 'No. Plots', 'Plot Wt.'),
               caption = "Original Sample Design for the Entire Sample Frame", 
               align = c("c", rep('r', times = (ncol(.) - 1)))) 

rm(r, plot_summary)
```

```{r Adjust the Sample Point Weights across entire Field Office}

Full_newWghts <- plotWeigher(OriginalWeights, pts)
Full_newWghts %>% 
  mutate(across(TotalAcres:WgtAcres, ~ round(.x))) %>% 
  mutate(across(PlotsSampled:PlotsRejected, ~ replace_na(.x, 0))) %>% 
  arrange(-TotalAcres) %>% 
  select(Stratum, PropInference, AreaInference:WgtAcres, PlotsSampled, PlotsRejected) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","),
               col.names = c('Stratum', 'Inf. Prop.', 
                             'Area Inf.', 'Plot Wt.', 'Sampled', 'Rejected'),
               caption = "Realized Weighted Sample Design for the Entire Sample Frame", 
               align = c("c", rep('r', times = (ncol(.) - 1))))

OriginalWeights_results <- OriginalWeights
NewWeights_results <- Full_newWghts

rm(OriginalWeights, Full_newWghts)
```

> "*greater than 70 percent of vegetation communities on (the remaining) BLM-administered lands*"

\newpage

The reporting units of Areas of Critical Environmental Concern (ACEC's), and Wilderness Study Areas (WSA), have different management objectives relative to the remaining BLM administered surface area. Loosely, they are intended to have a higher percentage of the vegetation communities within the natural range of variation: 

> "*greater than 80 percent of vegetation communities in ACEC's, WSA's...*" 

These areas were not intensified units within the original sample design, rather we split them out here using the original point draw for the field office. Here we calculate the initial sample weights for them using the same approach as for the remainder of BLM land, i.e. the acreage of each stratum is weighed against a targeted proportion of plots in the region. As our sample design was initiated and completed during a period of drought (See...), we dismiss the possibilities of making temporal comparisons across the sample panels. Accordingly, we have strata within these management units which: do not have a point per year panel (i.e. cannot be sampled each year). Subsequently, we do not have the initial ability to infer across the entire acreage of each stratum within them. 

```{r Original Point weights for ACEC and WSA}

spatial_products <- '/media/sagesteppe/ExternalHD/UFO_cartography'
design_rast <- '/media/sagesteppe/ExternalHD/UFO_AIM_Design_Stratification/processed'
lookup_table <- read.csv(file.path(design_rast, 'UFO_strata_areas.csv')) %>% 
  select(RasterValue, Code)

strat_raster <- terra::rast(file.path(design_rast, 'UFO_Strata_reclass_30m.tif'))
uFO <- st_read(file.path(spatial_products, 
                         'BLM_CO_Grazing_Allotments/gra_allot_poly.shp'), quiet = T)
acecs <- st_read(file.path(spatial_products, 'BLM_CO_ACEC/acec_desig_poly.shp'), 
                 quiet = T) %>%  
  st_intersection(uFO, .) %>% 
  select(NAME = ACEC_NAME) %>% 
  mutate(TYPE = 'ACEC')
wsa <- st_read(file.path(spatial_products, 'BLM_CO_WSA/nlcs_wsa_poly.shp'), 
               quiet = T) %>% 
  st_intersection(uFO, .) %>% 
  select(NAME = NLCS_NAME) %>% 
  mutate(TYPE = 'WSA')

wsAcec <- bind_rows(acecs, wsa) %>% 
  mutate(ID = 1:n(), .before = 1)

rm(acecs, wsa)

wsAcec_OriginalWeights <- originalWeights(sample_area = wsAcec, strat_raster = strat_raster, pt_data = pts)
wsAcec_pts <- wsAcec_OriginalWeights[[1]] # we return a list with the points needed for recalculating the weights
wsAcec_OriginalWeights <- wsAcec_OriginalWeights[[2]]

wsAcec_OriginalWeights %>% 
  mutate(PropArea = round(PropArea, 3)) %>% 
  mutate(across(.cols = c(Acres, ApproxStWgt),  round)) %>% 
  arrange(-Acres) %>% 
  ungroup() %>% 
  select(Stratum, Acres, PropArea, PropTarget, DesiredSS, ApproxStWgt) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","), 
               col.names = c('Stratum', 'Total Area (acres)', 'Prop. Area', 
                             'Prop. Site', 'No. Plots', 'Plot Wt.'),
               caption = "Original Sample Design for Areas of Critical Environmental Concern
               & Wilderness Study Areas", 
               align = c("c", rep('r', times = (ncol(.) - 1)))) 

rm(raster_cells, wsaSpat, wsAcec, wsAcec_area_summary, wsAcec_pt_draw,
   base_wsAcec_pts, raster_cells, wsAcec_areas, design_rast)
```

Strata with five or more plots, would allow for temporal analyses to be conducted on their data. Strata with less than five plots can only be treated as static entities within this time period 

```{r Final Point weights for ACEC and WSA}
wsAcec_newWghts <- plotWeigher(wsAcec_OriginalWeights, wsAcec_pts)

wsAcec_newWghts %>% 
  mutate(across(TotalAcres:WgtAcres, ~ round(.x))) %>% 
  mutate(across(PlotsSampled:PlotsRejected, ~ replace_na(.x, 0))) %>% 
  arrange(-TotalAcres) %>% 
  select(Stratum, PropInference, AreaInference:WgtAcres, PlotsSampled, PlotsRejected) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","),
               col.names = c('Stratum', 'Inf. Prop.',  
                             'Area Inf.', 'Plot Wt.', 'Sampled', 'Rejected'),
                             caption = "Realized Sample Design for Areas of Critical Environmental Concern
               & Wilderness Study Areas",               
               align = c("c", rep('r', times = (ncol(.) - 1))))

OriginalWeights_results <- bind_rows(OriginalWeights_results, wsAcec_OriginalWeights)
NewWeights_results <- bind_rows(NewWeights_results, wsAcec_newWghts)

rm(wsAcec_OriginalWeights, wsAcec_pts)
```

\newpage

```{r Original Point weights for Dominguez-Escalente NM}

monuments <- st_read(file.path(spatial_products, 'BLM_CO_NM_NCA/nlcs_nm_nca_poly.shp'), 
                 quiet = T) %>% 
  st_intersection(uFO, .) %>% 
  group_by(NLCS_NAME) %>% 
  st_collection_extract("POLYGON")  %>%  # there are these funny points
  summarize(geometry = st_union(geometry)) %>% 
  rename(NAME = NLCS_NAME)

DEnm <- monuments %>% filter(str_detect(NAME, 'Dominguez')) %>% 
  mutate(ID = 1:n(), .before = 1, TYPE = 'NM')

OW_deNM <- originalWeights(DEnm, strat_raster, pts)
OW_deNM_pts <- OW_deNM[[1]]
OW_deNM <- OW_deNM[[2]]

OW_deNM %>% 
  mutate(PropArea = round(PropArea, 3)) %>% 
  mutate(across(.cols = c(Acres, ApproxStWgt),  round)) %>% 
  arrange(-Acres) %>% 
  ungroup() %>% 
  select(Stratum, Acres, PropArea, PropTarget, DesiredSS, ApproxStWgt) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","), 
               col.names = c('Stratum', 'Total Area (acres)', 'Prop. Area', 
                             'Prop. Site', 'No. Plots', 'Plot Wt.'),
               caption = "Original Sample Design for Dominguez-Escalante National Monument", 
               align = c("c", rep('r', times = (ncol(.) - 1)))) 

rm(uFO, p2d, DEnm)
```

Strata with five or more plots, would allow for temporal analyses to be conducted on their data. Strata with less than five plots can only be treated as static entities within this time period 

```{r Final Point Weights for Dominguez-Escalente NCA}

OW_newWghts <- plotWeigher(OW_deNM, OW_deNM_pts)
OW_newWghts %>% 
  mutate(across(TotalAcres:WgtAcres, ~ round(.x))) %>% 
  mutate(across(PlotsSampled:PlotsRejected, ~ replace_na(.x, 0))) %>% 
  arrange(-TotalAcres) %>% 
  select(Stratum, PropInference, AreaInference:WgtAcres, PlotsSampled, PlotsRejected) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","),
               col.names = c('Stratum', 'Inf. Prop.', 
                             'Area Inf.', 'Plot Wt.', 'Sampled', 'Rejected'),
               caption = "Realized Sample Design for Dominguez-Escalante National Monument", 
               align = c("c", rep('r', times = (ncol(.) - 1))))

OriginalWeights_results <- bind_rows(OriginalWeights_results, OW_deNM)
NewWeights_results <- bind_rows(NewWeights_results, OW_newWghts)

rm(OW_deNM, OW_deNM_pts)
```

\newpage

Under the master sample design the number of plots which could be inferred from in the Salt Desert stratum would be 17. However, due to an intensification a surplus of 5, plots were sampled, however as these plots exceed the number of ... Weird problem. 

```{r Original Point Weights for Gunnison Gorge NCA}

GGnm <- monuments %>% filter(str_detect(NAME, 'Gunnison')) %>% 
  mutate(ID = 1:n(), .before = 1, TYPE = 'NM')
  
GG_deNM <- originalWeights(GGnm, strat_raster, pts)
GG_deNM_pts <- GG_deNM[[1]]
GG_deNM <- GG_deNM[[2]]

GG_deNM %>% 
  mutate(PropArea = round(PropArea, 3)) %>% 
  mutate(across(.cols = c(Acres, ApproxStWgt),  round)) %>% 
  arrange(-Acres) %>% 
  ungroup() %>% 
  select(Stratum, Acres, PropArea, PropTarget, DesiredSS, ApproxStWgt) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","), 
               col.names = c('Stratum', 'Total Area (acres)', 'Prop. Area', 
                             'Prop. Site', 'No. Plots', 'Plot Wt.'),
               caption = "Original Sample Design for Dominguez-Escalante National Monument", 
               align = c("c", rep('r', times = (ncol(.) - 1)))) 

rm(deSS, lookup_table, monuments, strat_raster, GGnm)
```

```{r Final Points Weights for Gunnison Gorge NCA}

GG_newWghts <- plotWeigher(GG_deNM, GG_deNM_pts)
GG_newWghts %>% 
  mutate(across(TotalAcres:WgtAcres, ~ round(.x))) %>% 
  mutate(across(PlotsSampled:PlotsRejected, ~ replace_na(.x, 0))) %>% 
  arrange(-TotalAcres) %>% 
  select(Stratum, PropInference, AreaInference:WgtAcres, PlotsSampled, PlotsRejected) %>% 
  knitr::kable(., digits = 2, format.args = list(big.mark = ","),
               col.names = c('Stratum', 'Inf. Prop.', 
                             'Area Inf.', 'Plot Wt.', 'Sampled', 'Rejected'),
               caption = "Realized Sample Design for Gunnison Gorge National Monument", 
               align = c("c", rep('r', times = (ncol(.) - 1))))

OriginalWeights_results <- bind_rows(OriginalWeights_results, GG_deNM)
NewWeights_results <- bind_rows(NewWeights_results, GG_newWghts)

rm(GG_deNM, GG_deNM_pts, pts)
```

\newpage

```{r Alluvial tracking plot status, warning = F, fig.height = 4, fig.weight = 4, fig.cap = 'Fates of all potential AIM plots from the Sample Design'}

ps <- plot_status %>% 
  filter(str_detect(Panel, 'OverSample', negate = T)) %>% 
  group_by(stratum, Plot.Status) %>% 
  count() %>% 
  mutate(across(where(is.numeric), ~ replace_na(.x, 0))) %>% 
  drop_na() %>% 
  bind_rows(., data.frame(
    stratum = 'GR', Plot.Status = 'rejected', n = 0)) %>% 
  mutate(Plot.Status = factor(Plot.Status, levels = c('sampled', 'rejected')),
         stratum = factor(stratum, levels = c('AS', 'SS', 'GR', 'MMS', 'OT', 'PJ', 'PP', 'SD', 'RI', 'MC')))

ggplot(data = ps,
       aes(axis1 = stratum, axis2 = Plot.Status, y = n), color = NA) +
  geom_alluvium(aes(fill = Plot.Status), curve_type = "arctangent", width = 0.25) +
  geom_stratum(color = 'grey85') +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), color = 'grey15') +
  scale_fill_manual(values = c('forestgreen', '#e71313')) +
  theme_void() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("Fate of Potential Plots Across the Sample Design") 

rm(ps)
```

```{r}

```

